---
title: "Homework 2"
author: Logan Caraco
output: 
  html_document:
    toc: true
    toc_float: true
---

``` {r prelude}
library(ggplot2)
library(dplyr)
library(fivethirtyeight)

library(magrittr)

data("bechdel")
```


\
\



# Exercise 1

## Part a
``` {r process-bechdel}
# Define a new variable which measures years_since_1900
bechdel %<>% mutate(years_since_1900 = year - 1900)

# NOTE: var-name %<>% func is syn sugar for var-name <- var-name %>% func
```

```{r peek-bechdel}
bechdel %>% 
  select(year, years_since_1900) %>%
  head(3)
```

## Part b

``` {r add-mils}
bechdel %<>% 
  mutate(budget_mil = budget / 1e6) %>%
  mutate(intgross_mil = intgross / 1e6)
```


``` {r peek-mils}
bechdel %>% 
  select(budget_mil, intgross_mil) %>% 
  head(3)
```

\
\


# Exercise 2

``` {r}
# How many films are in the dataset?
nrow(bechdel)

# When was the oldest film in the dataset made?
bechdel %>%
  summarize(min(year))

# When was the most recent film in the dataset made?
bechdel %>%
  summarize(max(year))

# What is the median film budget in 2013 dollars?
bechdel %>% 
  summarize(median(budget_2013))

# What films had budgets that exceeded 300 million in 2013 dollars?
# (Just show the films and budgets, not all variables)
bechdel %>%
  select(title, budget_2013) %>%
  filter(budget_2013 > 300000000)

# What films were made in 1971?
# (Just show the films and year, not all variables)
bechdel %>%
  select(title, year) %>%
  filter(year == 1971)

# What were the budget_mil and intgross_mil values for "Titanic"?
# (Just show the budget_mil and intgross_mil values)
bechdel %>%
  filter(title == "Titanic") %>%
  select(budget_mil, intgross_mil)
```

# Exercise 3

``` {r}
bechdel %>% 
  group_by(clean_test) %>% 
  summarize(count = n())

ggplot(bechdel, aes(x = clean_test)) + 
  geom_histogram(color = "white", stat="count")
```


\
\


# Exercise 4

The graph below presents the distribution of total international gross (in millions) for movies from 1970-2013 (adjusted to 2013 dollars where necessary). 

The vast majority of movies make less than 200 million dollars, but there are movies that made in excess of $1B. 
``` {r}
ggplot(bechdel, aes(x = intgross_mil)) +
  geom_histogram(color = "white", binwidth = 25) + 
  xlim(-50, 1250)
```

``` {r}
bechdel %>%
  summarize(mean(intgross_mil, na.rm=TRUE), median(intgross_mil, na.rm=TRUE))
```

The median is much lower than the mean because there is a long and short right tail which contributes heavily to the median, and barely at all to the median.

``` {r}
bechdel %>%
  summarize(sd(intgross_mil, na.rm=TRUE), sd(budget, na.rm=TRUE)/1000000)
```
The budget is relatively tightly grouped, whereas the gross is highly variable. Therefore it is a risky venture to create a high budget movie, because the chance of payout can be slim.


# Exercise 5

$\texttt{intgross_mil}$ is the response variable to $\texttt{budget}$.

```{r}
bechdel %>%
  ggplot(aes(y = intgross_mil, x = budget_mil)) + 
  geom_point() +                         # <<< plot
  geom_smooth(method = "lm", se = FALSE) # <<< trend
```


As seen in the plot above, there is an (unsuprisingly) positive relationship between budget and international gross, thought the underestimates of the model tend to be greater than the overestimates.

# Exercise 6

``` {r}
summary(movie_model <- lm(intgross_mil ~ budget_mil, data = bechdel))$coefficients
```

The regression trendline:
\[\texttt{intgross_mil} = 11.95 + 3.07~ \texttt{budget_mil}\]
Roughly, every million dollars put into the budget of the movie increases the international gross by three million dollars (or return on a \$1M investment is twofold).

The intercept is not totally intuitive as a movie with budget of \$0 would unlikely have a \$12M gross, it makes more sense to say that million-dollar budget movies tend to gross 15M.

# Exercise 7

\[\texttt{budget_mil}_{Titanic} = 200\]
\[\texttt{intgross_mil} = 11.95 + 3.07~ \texttt{budget_mil} = 625.95\]

``` {r}
predict(movie_model, newdata = data.frame(budget_mil = 200))
```

\[\texttt{intgross_mil}_{actual}=2185.672\]
\[\textbf r = actual - predicted = 2185.672 - 626.572 = 1559.1\]

The model *grossly* underestimated the gross.

# Exercise 8

``` {r}
ggplot(bechdel, aes(x = intgross_mil, color = clean_test)) +
  geom_histogram(color = "white", binwidth = 50) + 
  facet_wrap(~ clean_test) +
  xlim(-50, 1250)
```


``` {r}
bechdel %>%
  group_by(clean_test) %>%
  summarize(mean(intgross_mil, na.rm=TRUE), sd(intgross_mil, na.rm=TRUE))
```

# Exercise 9

``` {r}
bechdel <- bechdel %>% 
  mutate(clean_test = factor(as.character(clean_test)))
movie_model_2 <- lm(intgross_mil ~ clean_test, data = bechdel)
summary(movie_model_2)$coefficients
```

The baseline for $\texttt{clean_test}$ is $\texttt{dubious}$.

$\texttt{dubious}$ movies tend to gross 169 million. $\texttt{men}$ movies tend to make 23M less than $\texttt{dubious}$ ones. 

The $\texttt{notalk}$ movies (in which women are present on screen but do not speak) have the highest average gross of 175M. The $\texttt{ok}$ movies (which pass the Bechdel test) have the lowest average gross of 132.7M.
