---
title: "Homework 5"
author: Logan Caraco
output: 
  html_document:
    toc: true
    toc_float: true
---


\
\

``` {r, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)

# Load the package
library(fivethirtyeight)

# Load the data
data("candy_rankings")
```


# Exercise 1

```{r}
# What's the average winpercent across all candies?
candy_rankings %>%
  select(winpercent) %>%
  summarize(mean(winpercent))

# What's the average winpercent among chocolate vs non-chocolate candies?
candy_rankings %>%
  group_by(chocolate) %>%
  select(winpercent) %>%
  summarize(mean(winpercent))

# What's the average winpercent among candies that have both chocolate & caramel?
candy_rankings %>%
  filter(chocolate, caramel) %>%
  select(winpercent) %>%
  summarize(mean(winpercent))

# What are the 6 most popular candies? 
# (Show just the candy names & winpercent)
candy_rankings %>%
  arrange(desc(winpercent)) %>%
  select(competitorname, winpercent) %>%
  head(6)

# What are the 6 least popular candies? 
# (Show just the candy names & winpercent)
candy_rankings %>%
  arrange(winpercent) %>%
  select(competitorname, winpercent) %>%
  head(6)

# What are the 6 most popular fruity candies?
# (Show just the candy names & winpercent)
candy_rankings %>%
  filter(fruity) %>%
  arrange(desc(winpercent)) %>%
  select(competitorname, winpercent) %>%
  head(6)
```

\
\


# Exercise 2

## Part a

``` {r}
candy_rankings %>%
  ggplot(aes(x = fruity, y = winpercent, color = hard)) +
  geom_boxplot()
```

## Part b

``` {r}
candy_model_1 <- lm(winpercent ~ fruity + hard, candy_rankings)
summary(candy_model_1)$coefficients
```

## Part c

\[-9.015980 \pm 2 * 3.203994 \approx (-15.4, -2.6)\]

```{r}
confint(candy_model_1)
```

## Part d
It is very likely that fruitiness decreases odds of winning a match-up in the model.

## Part e
What conclusion can we make from this CI: when controlling for whether or not a candy is hard…
**fruity candy tends to be significantly less popular than non-fruity candy.**


# Exercise 3

```{r}
candy_rankings %<>% 
  mutate(pricepercent = 100*pricepercent)
```

## Part a
```{r}
candy_rankings %>%
  ggplot(aes(x = pricepercent, y = winpercent)) +
  geom_jitter() +
  ylim(0, 100)
```

## Part b
Probably not, there appears to be an almost perfectly horizontal line.

## Part c
Probably not, there appear to be a full array of `winpercent` at each cluster of `pricepercent`.


# Exercise 4

## Part a
```{r}
candy_model_2 <- lm(winpercent ~ pricepercent, candy_rankings)
summary(candy_model_2)$coefficients
```

## Part b

```{r}
summary(candy_model_2)$r.squared
```

With such a low r-squared, this is a terrible model.

## Part c

```{r}
candy_rankings %>%
  select(competitorname, pricepercent, winpercent) %>%
  # arrange first by distance from median price, then by popularity
  arrange(abs(pricepercent - 50), desc(winpercent)) %>%
  head(3)

predict(candy_model_2, data.frame(pricepercent = 51.1), interval="predict") 
```

# Exercise 5
## Part a
\[H_0: \beta_1 = 0\]
\[H_a: \beta_1 > 0\]

## Part b
```{r}
summary(candy_model_2)$coefficients
(0.1778277 - 0) / 0.05304668
```

## Part c
```{r}
1.208829e-03 / 2
```

It is unlikely that this positive correlation was observed by chance (much less than 0.1%).

## Part d
We reject the null hypothesis. There appears to be a very slight positive correlation between price and popularity.
